---
title: "DE genes intersects and comparisons"
author: "Elsie Jacobson"
date: "4/20/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(Vennerable)
library(extrafont)
library(ggrepel)

ggformat <- theme_classic(base_family = "Arial") + theme(text = element_text(size = 20))
```

###Prepare datasets

Read all DGE datasets

```{r}
mig5v14 <- read.csv('../results/mig5v14_all.csv', row.names=1)
mig5v14_sig <- read.csv('../results/mig5v14_sig.csv', row.names=1)
mig5 <- read.csv('../results/mig5_all.csv', row.names=1)
mig5_sig <- read.csv('../results/mig5_sig.csv', row.names=1)
mig14 <- read.csv('../results/mig14_all.csv', row.names=1)
mig14_sig <- read.csv('../results/mig14_sig.csv', row.names=1)
```

Volcano plots

```{r}
ggplot(mig5v14, aes(log2FoldChange, -log(padj), colour=padj<0.05)) + geom_point(alpha=0.5) + theme_bw() + scale_colour_manual(values=c("#2F4F4F", "red"))
ggsave("../plots/volcano_5v14.pdf", useDingbats=FALSE)

ggplot(mig5, aes(log2FoldChange, -log(pvalue), colour=padj<0.05)) + geom_point(alpha=0.5) + theme_bw() + scale_colour_manual(values=c("#2F4F4F", "red"))
ggsave("../plots/volcano_5vctrl.pdf", useDingbats=FALSE)

ggplot(mig14, aes(log2FoldChange, -log(pvalue), colour=padj<0.05)) + geom_point(alpha=0.5) + theme_bw() + scale_colour_manual(values=c("#2F4F4F", "red"))
ggsave("../plots/volcano_14vctrl.pdf", useDingbats=FALSE)
```

Compile dataset of DGE in 3 comparisons

```{r}
hgnc_ens <- readRDS('../data/hgnc_ens.rds')
convert <- hgnc_ens %>%
  dplyr::select(hgnc_symbol, ensembl_gene_id) %>%
  distinct()

n5v14 <- paste(colnames(mig5v14), "5v14",sep="_")
n5 <- paste(colnames(mig5), "5",sep="_")
n14 <- paste(colnames(mig14), "14",sep="_")

colnames(mig5v14) <- n5v14
colnames(mig5) <- n5
colnames(mig14) <- n14

allmig <- mig5v14 %>%
  full_join(mig5, by=c("gene_5v14"="gene_5")) %>%
  full_join(mig14, by=c("gene_5v14"="gene_14"))

allmig_sig <- allmig %>%
  left_join(convert, by=c("ens_gene_14"="ensembl_gene_id")) %>%
  mutate(sig_5v14=ifelse(padj_5v14<0.05,TRUE,FALSE)) %>%
  mutate(sig_5=ifelse(padj_5<0.05,TRUE,FALSE)) %>%
  mutate(sig_14=ifelse(padj_14<0.05,TRUE,FALSE)) %>%
  filter(sig_5v14==TRUE | sig_5==TRUE | sig_14==TRUE) %>%
  mutate(set=case_when(sig_5==TRUE & sig_14==TRUE~'set1',
                       sig_5==TRUE & sig_14==FALSE & sig_5v14==FALSE~'set2',
                       sig_5v14==TRUE & sig_14==FALSE~'set3')) %>%
  dplyr::select(hgnc_symbol, ext_gene=external_gene_name_14, ens_gene_id=ens_gene_14,log2FoldChange_5v14,padj_5v14,log2FoldChange_5vctrl=log2FoldChange_5,padj_5vctrl=padj_5,
                log2FoldChange_14vctrl=log2FoldChange_14,padj_14vctrl=padj_14, set) %>%
  na.omit()

write.csv(allmig_sig, "../results/Significant_DE_genes_all_conditions.csv")
```

###Venn diagram of the intersects

```{r}
venrbl <- NULL
venrbl$mig5v14 <- as.character(mig5v14_sig$ens_gene)
venrbl$mig5 <- as.character(mig5_sig$ens_gene)
venrbl$mig14 <- as.character(mig14_sig$ens_gene)
str(venrbl)

Vmig <- Venn(venrbl)
Vmig
pdf("../plots/migVenn.pdf")
plot(Vmig,doWeights=FALSE, show = list(SetNames=c("5 vs 14","none vs 5", "none vs 14"), Faces = FALSE, DarkMatter = FALSE))
dev.off()
```

###Compare changes between mig5 vs control and mig14 vs control in different subsets

Significant in both mig5 vs control and mig14 vs control

```{r}
mod <- lm(log2FoldChange_14~log2FoldChange_5, data=subset(allmig_sig, sig_5==TRUE & sig_14==TRUE))
summary(mod)
ggplot(subset(allmig_sig, sig_5==TRUE & sig_14==TRUE), aes(log2FoldChange_5, log2FoldChange_14)) + geom_point(alpha=0.5) + geom_abline(slope=0.78536, intercept = -0.01094) + ggformat + labs(x=expression(paste("5",mu,"m vs control logFC")), y=expression(paste("14",mu,"m vs control logFC")))
ggsave("../plots/logFC_5_14_significant.pdf")
```

Significant only in mig5 vs control

```{r}
mod <- lm(log2FoldChange_14~log2FoldChange_5, data=subset(allmig_sig, sig_5==TRUE & sig_5v14==FALSE & sig_14==FALSE))
summary(mod)
ggplot(subset(allmig_sig, sig_5==TRUE & sig_5v14==FALSE & sig_14==FALSE), aes(log2FoldChange_5, log2FoldChange_14)) + geom_point(alpha=0.5) + geom_abline(slope=0.473765, intercept = 0.007815) + ggformat + labs(x=expression(paste("5",mu,"m vs control logFC")), y=expression(paste("14",mu,"m vs control logFC")))
ggsave("../plots/logFC_5_only_significant.pdf")
```

Significant in mig5 vs mig14 but not in mig14 vs control - remodelling only!

```{r}
mod <- lm(log2FoldChange_14~log2FoldChange_5, data=subset(allmig_sig, sig_14==FALSE & sig_5v14==TRUE))
summary(mod)
ggplot(subset(allmig_sig, sig_14==FALSE & sig_5v14==TRUE), aes(log2FoldChange_5, log2FoldChange_14)) + geom_point(alpha=0.5) + ggformat + labs(x=expression(paste("5",mu,"m vs control logFC")), y=expression(paste("14",mu,"m vs control")))
ggsave("../plots/logFC_5v14sig_mig14nonsig.pdf")
```

###Convert mig5 vs mig14 results to bed format for further analysis

```{r}
mig5v14_bed <- mig5v14 %>%
  left_join(g2o_full, by=c("ens_gene"="ensembl_gene_id")) %>%
  mutate(chromosome_name=paste("chr",chromosome_name,sep="")) %>%
  mutate(chromosome_name=gsub("chrMT","chrM",chromosome_name)) %>%
  mutate(strand=ifelse(strand>0,'+','-')) %>%
  dplyr::select(chromosome_name, start_position, end_position, ens_gene,strand) %>%
  distinct()

mig5v14_sig_bed <- mig5v14_sig %>%
  left_join(g2o_full, by=c("ens_gene"="ensembl_gene_id")) %>%
  mutate(chromosome_name=paste("chr",chromosome_name,sep="")) %>%
  mutate(chromosome_name=gsub("chrMT","chrM",chromosome_name)) %>%
  mutate(strand=ifelse(strand>0,'+','-')) %>%
  dplyr::select(chromosome_name, start_position, end_position, ens_gene,strand) %>%
  distinct()

write.table(mig5v14_sig_bed, '../results/mig5v14_sig.bed', sep='\t', quote=FALSE, row.names = FALSE, col.names = FALSE)
write.table(mig5v14_bed, '../results/mig5v14.bed', sep='\t', quote=FALSE, row.names = FALSE, col.names = FALSE)
```