---
title: "tiled_interactions"
author: "Elsie Jacobson"
date: "7/10/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(fuzzyjoin)
library(scales)
#library(plyr)
#detach("package:plyr")

colsame <- viridis_pal(option="B", begin=0.2, end=0.8)(2)
```

```{r}
up5 <- read.csv("../data/5migv_ctrl_100_40_ints.txt", sep="\t", stringsAsFactors = FALSE)
up14 <- read.csv("../data/14migv_ctrl_100_40_ints.txt", sep="\t", stringsAsFactors = FALSE)
down5 <- read.csv("../data/ctrl_v_5mig_100_40_ints.txt", sep="\t", stringsAsFactors = FALSE)
down14 <- read.csv("../data/ctrl_v_14mig_100_40_ints.txt", sep="\t", stringsAsFactors = FALSE)
up5v14 <- read.csv("../data/5mig_v_14mig_100_40_ds_ints.txt", sep="\t", stringsAsFactors = FALSE)
down5v14 <- read.csv("../data/14mig_v_5mig_100_40_ds_ints.txt", sep="\t", stringsAsFactors = FALSE)
```

###Function to filter significant and differential contacts from a homer output file

```{r}
make_sig_diff <- function(data, out_sig, out_diff){
  raw <- read.csv(data, sep = '\t', stringsAsFactors = FALSE)
  dat <- raw %>%
    filter(chr.1.==chr.2.,chr.1.!="chrY"& chr.1.!="chrM")
    
  sig <- dat %>% 
    mutate(dist=as.numeric(Distance)) %>%
    filter(dist>=100000) %>%
    filter(FDR.Benjamini..based.on.2.98e.09.total.tests. <= 0.05) %>%
    mutate(pval=2*pnorm(-abs(Z.score.Difference.vs..Bg))) %>%
    mutate(qval=p.adjust(pval, method = "BH"))
  
  diff <- sig %>%
    filter(qval <= 0.1)
  
  write.csv(sig, out_sig, row.names=FALSE)
  write.csv(diff, out_diff, row.names=FALSE)
}
```

##Mig5 vs control 

###Increased contacts

```{r}
make_sig_diff('../data/5migv_ctrl_100_40_ints.txt', "../results/mig5_significant_ints.csv", "../results/mig5_up_filtered.csv")
```

###Decreased contacts

```{r}
make_sig_diff('../data/ctrl_v_5mig_100_40_ints.txt', "../results/ctrl_significant_ints.csv", "../results/mig5_down_filtered.csv")
```

##Mig14 vs control 

###Increased contacts

```{r}
make_sig_diff('../data/14migv_ctrl_100_40_ints.txt', "../results/mig14_significant_ints.csv", "../results/mig14_up_filtered.csv")
```

###Decreased contacts

```{r}
make_sig_diff('../data/ctrl_v_14mig_100_40_ints.txt', '../results/ctrl_significant_ints_v14.csv', "../results/mig14_down_filtered.csv")
```

##Mig5 vs mig14 

###Increased contacts

```{r}
make_sig_diff('../data/5mig_v_14mig_100_40_ds_ints.txt', '../results/mig5_significant_ints_v14.csv', "../results/mig5v14_up_filtered.csv")
```

###Decreased contacts

```{r}
make_sig_diff('../data/14mig_v_5mig_100_40_ds_ints.txt', '../results/mig14_significant_ints_v5.csv', "../results/mig5v14_down_filtered.csv")
```

##Function to convert filtered homer files to bed format

```{r}
homer2bed <- function(x) {

  dat <- read.csv(x)
  filename=ifelse(str_detect(x,"significant"),gsub("significant_ints.csv","sig.bed",x),
                  ifelse(str_detect(x,"filtered"),gsub("filtered.csv","diff.bed",x), "WRONG FILE"))
  
  r2 <- dat %>%
    dplyr::select(chrom=chr.2., 
                  start=start.2., end=end.2., 
                  name=InteractionID, Distance) %>%
    mutate(start=ifelse(start<0, 0, start)) %>%
    mutate(end=ifelse(end<0, 0, end)) %>%
    distinct()
  
  bed <- dat %>%
    dplyr::select(chrom=chr.1., 
                  start=start.1., end=end.1., 
                  name=InteractionID, Distance) %>%
    mutate(start=ifelse(start<0, 0, start)) %>%
    mutate(end=ifelse(end<0, 0, end)) %>%
    distinct() %>%
    bind_rows(r2)
  
  write.table(bed, filename, sep="\t", col.names = FALSE ,row.names = FALSE, quote=FALSE)
  return(filename)
  
}
```

```{r}
homer2bed("../results/mig5_significant_ints.csv")
homer2bed("../results/ctrl_significant_ints.csv")
homer2bed("../results/mig14_significant_ints.csv")

homer2bed("../results/mig5_up_filtered.csv")
homer2bed("../results/mig5_down_filtered.csv")
homer2bed("../results/mig14_up_filtered.csv")
homer2bed("../results/mig14_down_filtered.csv")
homer2bed("../results/mig5v14_up_filtered.csv")
homer2bed("../results/mig5v14_down_filtered.csv")
```

add PC1 values with bedtools 'add_PC1.zsh'

move to tiled_interaction_analysis.Rmd